The Socialization of Meritocracy: The role of the Family and the School

### 2023-06-15 docs/presentacion-effort/presentation-effort.Rmd
ajustes presentacion

---
title: "The Socialization of Meritocracy: The role of the Family and the School"
subtitle: ".medium[**Resultados preliminares**]"
author: "Juan Carlos Castillo, Julio Iturra, Francisco Meneses & Martín Venegas"
institute: ".small[**FONDECYT N°1210847 Meritocracia en la escuela**]"
output:
  xaringan::moon_reader:
    css: ["custom_lisa.css"]
    seal: false
    nature:
      highlightStyle: solarized-dark
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "xaringan_custom/macros.js"
      ratio: '16:9'
---

@@

----
<br>
.espaciosimplelineas[
.medium[
Juan Carlos Castillo, Julio Iturra, Francisco Meneses & Martín Venegas
] 
]

@@

layout: true
class: animated, fadeIn

---
class: middle

.pull-left-narrow[
# **PRICE**:
### Programa de Investigación en Ciudadanía Escolar
]

.pull-right-wide[
![](images/price.png)

.right[
[ciudadania-escolar.cl/](https://ciudadania-escolar.cl)
]]

---
class: middle

.pull-left-narrow[
# **PRICE**:
### Programa de Investigación en Ciudadanía Escolar
]

.pull-right-wide[

![](images/proyectos.png)

.right[
Projects since 2011 ]
]

---
.pull-left-narrow[
# **PRICE**:
### Programa de Investigación en Ciudadanía Escolar

(School Citizenship Research Programme)
]

.pull-right-wide[

![](images/equipo.png)
]

@@

# .black[Introduction]

.pull-left[
  .content-box-gray[
## Meritocracy
  - Effort and talent  
  - Descriptive (perception) 
  - Normative (preferences)   
  ] 
]

--
.pull-right[
<br><br>

.content-box-green[
## Socialisation
  - Family and school as socialisation agents 
  - Justice experiences at school
  ]  
]

###.yellow[
_Is there an association between the meritocratic perceptions of parents with those of their offsprings?_
]

### .yellow[_To what extent is school experience in terms of distributive justice (in relation to grades) linked to the meritocratic perceptions of students?_]

---
# .black[Introduction]

----

@@

  - Socio-economic status (self-interest)

  - Intergenerational transmission of beleifs and attitudes (socialisation)

@@

.pull-right[
.content-box-yellow[
## School

- Meritocracy perceptions (school and society)

- Justice evaluations about grades

@@

@@

* Students, parents and teachers

* Analytical sample ( $n_i$ =541)

  - Student and parent information
  - $n_j$ = 60 (schools)

* _Method:_  Multilevel Ordinal logit  
]

@@

### DV: Perceived Meritocracy

.center[![:scale 78%](images/perc-merit.jpg)]

---
### IV: Justice in grades

@@

*Indirect*

$$\text{Justice}= \ln\biggl(\frac{\text{grade obtained}}{\text{grade deserved}}\biggl)$$

.center[![:scale 93%](images/indirect-justice.png)]

@@

*Direct*

"Taking into account the time I devote to my studies, the grades I get are:"

.center[![:scale 100%](images/direct.png)]

@@

@@

.pull-left-narrow[
<br>
### In Chile those who work hard get ahead
]

.pull-right-wide[

.center[![:scale 85%](images/ploteffort.png)]
]

.pull-left-narrow[
<br>
### In Chile those who work hard get ahead
]

.pull-right-wide[
.center[![:scale 85%](images/intereffort.png)]
]

@@

- Family socialisation: mixed, relevant only for the importance of hard work
- SES-family has an inconsistent association. SES at the school level positively influences the importance of hard work but is null for perceived rewards.

- Justice experience in grades .bold[does not] have a direct influence on perceived meritocracy

- Interactions are mixed, decreasing the influence of socialisation on the importance of hard work, but positive effort is rewarded.

@@

----
<br>
.espaciosimplelineas[
.medium[
Juan Carlos Castillo, Julio Iturra, Francisco Meneses & Martín Venegas
] 
]

@@

>>>>

---
title: "The Socialization of Meritocracy: The role of the Family and the School"
subtitle: ".medium[**Resultados preliminares**]"
author: "Julio Iturra, Juan Carlos Castillo, Francisco Meneses & Martín Venegas"
institute: ".small[**FONDECYT N°1210847 Meritocracia en la escuela**]"
output:
  xaringan::moon_reader:
    css: ["custom_lisa.css"]
    seal: false
    nature:
      highlightStyle: solarized-dark
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "xaringan_custom/macros.js"
      ratio: '16:9'
---

@@

----
<br>
.espaciosimplelineas[
.small[
Julio Iturra -  Bremen Int. Graduate School of Social Sciences
<br>
Juan Carlos Castillo -  Universidad de Chile
<br>
Francisco Meneses & Martín Venegas
] 
]

@@

@@

---
<br>
.pull-left[
  .content-box-gray[
## Meritocracy
  - Merit: .bold[Effort] and talent (Young, 1958)  
  - Subjective meritocracy (Duru-Bellat & Tenret, 2012;Castillo et al, 2019)
    - Normative (_How it should be_) 
    - **Perception** (_How is it_) 
  ] 
]

.pull-right[
  .content-box-purple[
## Justice experiences
  - Justice experiences in school (Jasso & Resh, 2002)
  - Deserved vs obtained .bold[rewards]  (Resh & Sabbagh, 2014)
  ] 
]

---
# .black[Socialization agents]
----

@@

  - Self-interest: family socio-economic resources

  - Intergenerational transmission of beliefs and attitudes (Bandura et al., 2001; Olivos, 2021)

@@

.pull-right[
.content-box-yellow[

## School

- Meritocracy perceptions (Lampert, 2013; Mijs, 2018).

- Direct and indirect evaluation about grades (Resh, 2018; Resh & Sabbagh, 2014)

@@

###.yellow[.small[_What is the association between parental perceived meritocracy with their children's perceived meritocracy?_]]

### .yellow[.small[_To what extent the experiences of justice in grades influence student's perceived meritocracy?_]]

@@

- Highschool students (2nd) and parents (n = 541) nested in schools (n = 60)

* _Method:_  Multilevel regression models   
]

@@

### Perceived Meritocracy

_Today in Chile, for getting ahead in life, how important is (...) Hard work_

.center[![:scale 100%](images/perc-merit.jpg)]

---
###  Justice in grades: Indirect

@@

.bold[Obtained:] _What was your last year average grade? If you don't remember exactly, please give an approximate grade._

.bold[Deserved:] _And, what grade average do you think you deserved?_ 
$$\text{Justice}= \ln\biggl(\frac{\text{grade obtained}}{\text{grade deserved}}\biggl)$$

@@

.center[![:scale 100%](images/indirect-justice.png)]

@@

---
### Justice in grades:  Direct

_Taking into account the time I ddedicate to my studies, the grades I get are:_

.center[![:scale 90%](images/direct.png)]
  ]

@@

@@

- Family socialisation: parents influence student's perceived meritocracy

- SES-family has an inconsistent association. SES at the school does not influence student's perceived meritocracy in terms of the importance of hard work
- Justice experience in grades .bold[does not] have a direct influence on perceived meritocracy

- Negative justice experiences in grades descrease the influence of socialisation on the importance of hard work.

@@

----
<br>
.espaciosimplelineas[
.medium[
Julio Iturra -  Bremen Int. Graduate School of Social Sciences
<br>
Juan Carlos Castillo -  Universidad de Chile
<br>
Francisco Meneses & Martín Venegas
] 
]

@@

![:scale 35%](images/descriptivos1.png)

### 2023-06-15 paper-blinded.Rmd
ajustes presentacion

```{r plotgetah, eval=FALSE, include=FALSE}
# Funciones para renderizar tablas tanto en HTML como en pdf
if (!require("pacman")) install.packages("pacman")  #si falta pacman, instalar
if (!require("tinytex")) install.packages("tinytex")#si falta tinytex, instalar
pacman::p_load(knitr, kableExtra, dplyr, haven, texreg, sjPlot, lme4, ordinal, webshot) # librerias
knitr::opts_chunk$set(warning = FALSE,  # mensaje de warning
                      message = FALSE,  # mensajes/avisos de librerias
                      cache = FALSE,    # cache de los chunks,usar analisis pesados
                      out.width = '85%',# largo de imagen en %
                      fig.pos= "H",     # posicion figuras H = HERE
                      echo = FALSE      # incluir chunk en output
                      )

@@

getah.or6 <- ordinal::clmm(formula(paste("perc_trabajo_duro_factor_est~",paste(c("perc_tduro_bi_ap","sj_indirect_ln","sj_direct_factor",con,ref),collapse="+"))), basereg, nAGQ = 10)

getah.or8 <- ordinal::clmm(formula(paste("perc_trabajo_duro_factor_est~",paste(c("perc_tduro_bi_ap","sj_indirect_ln","sj_direct_factor", con, "perc_tduro_bi_ap*sj_direct_factor","nse_grupo",ref),collapse="+"))), basereg)

rm_plot <- c(
  "Not important|Somewhat important",
  "Somewhat important|Important",
  "Important|Very important",
  "quintiles_ingresos_pc_factorDon't know/No response",
  "libros_hogar_factor_apBetween 11 and 25 books",
  "libros_hogar_factor_apBetween 26 and 100 book",
  "libros_hogar_factor_apBetween 101 and 200 books",
  "libros_hogar_factor_apBetween 201 and 500 books",
  "libros_hogar_factor_apMore than 500 books",
  # "univ",
  "pos_pol_factor_apCenter Right",
  "pos_pol_factor_apCenter",
  "pos_pol_factor_apCenter Left",
  "pos_pol_factor_apLeft",
  "pos_pol_factor_apIndependent",
  "pos_pol_factor_apNone",
  "pos_pol_factor_apNa/Nr",
  "pos_pol_factor_estCenter Right",
  "pos_pol_factor_estCenter",
  "pos_pol_factor_estCenter Left",
  "pos_pol_factor_estLeft",
  "pos_pol_factor_estIndependent",
  "pos_pol_factor_estNone",
  "pos_pol_factor_estNa/Nr",
  "genero_factor_estMujer",
  "genero_factor_estOtro",
  "dependencia_factor_apPart. Subvencionado",
  "dependencia_factor_apPart. Privado")

@@

# plotgetah<- 
# plot_model(getah.or8,rm.terms = rm_plot,show.intercept = F,
#            show.p = T, show.values = T,
#            # order.terms = c(2,3,4,1,5,6,7,8,9,10,11,12),
#            axis.labels = rev(coef_names),
#            ) +
#   geom_hline(yintercept = 1)
# plotgetah
plotgetah<- 
plot_models(getah.or6,getah.or8,rm.terms = rm_plot,
            axis.labels = rev(coef_names),show.values = T,
            m.labels = c("Individual", "Interactions"),
            legend.title = "Models") +
  geom_hline(yintercept = 1) +
  theme(legend.position = "bottom")

ggplot2::ggsave(plot = plotgetah,device = "png",units = "cm",
                width = 1,height = 1,scale = 18,
                path =  "presentacion-effort/images/",
                filename = "plotgetahead.png")
# Elaborar modelo de interacción
getah.or7 <- ordinal::clmm(formula(paste("perc_trabajo_duro_factor_est~",paste(c("perc_tduro_bi_ap","sj_indirect_noln","sj_direct_factor", con, "perc_tduro_bi_ap*sj_direct_factor",ref),collapse="+"))), basereg)

ggplot2::ggsave(plot = graph_interact,device = "png",units = "cm",
                width = 1,height = 1,scale = 18,
                path =  "presentacion-effort/images/",
                filename = "intergetahead.png")
```

@@

effort.or6 <- ordinal::clmm(formula(paste("perc_esfuerzo_factor_est~",paste(c("perc_esfuerzo_bi_ap","sj_indirect_ln","sj_direct_factor", con,ref),collapse="+"))), basereg)
effort.or8 <- ordinal::clmm(formula(paste("perc_esfuerzo_factor_est~",paste(c("perc_esfuerzo_bi_ap","sj_indirect_ln","sj_direct_factor", con, "perc_esfuerzo_bi_ap*sj_direct_factor","nse_grupo",ref),collapse="+"))), basereg)

rm_plot <- c(
  "Strongly Disagree|Disagree",
  "Disagree|Agree",
  "Agree|Strongly Agree",
  "quintiles_ingresos_pc_factorDon't know/No response",
  "libros_hogar_factor_apBetween 11 and 25 books",
  "libros_hogar_factor_apBetween 26 and 100 book",
  "libros_hogar_factor_apBetween 101 and 200 books",
  "libros_hogar_factor_apBetween 201 and 500 books",
  "libros_hogar_factor_apMore than 500 books",
  # "univ",
  "pos_pol_factor_apCenter Right",
  "pos_pol_factor_apCenter",
  "pos_pol_factor_apCenter Left",
  "pos_pol_factor_apLeft",
  "pos_pol_factor_apIndependent",
  "pos_pol_factor_apNone",
  "pos_pol_factor_apNa/Nr",
  "pos_pol_factor_estCenter Right",
  "pos_pol_factor_estCenter",
  "pos_pol_factor_estCenter Left",
  "pos_pol_factor_estLeft",
  "pos_pol_factor_estIndependent",
  "pos_pol_factor_estNone",
  "pos_pol_factor_estNa/Nr",
  "genero_factor_estMujer",
  "genero_factor_estOtro",
  "dependencia_factor_apPart. Subvencionado",
  "dependencia_factor_apPart. Privado")

>>>>

```{r plotgetah, eval=FALSE, include=FALSE}
# Funciones para renderizar tablas tanto en HTML como en pdf
if (!require("pacman")) install.packages("pacman")  #si falta pacman, instalar
if (!require("tinytex")) install.packages("tinytex")#si falta tinytex, instalar
pacman::p_load(knitr, kableExtra, dplyr, haven, texreg, sjPlot, lme4, ordinal, webshot,ggplot2) # librerias
knitr::opts_chunk$set(warning = FALSE,  # mensaje de warning
                      message = FALSE,  # mensajes/avisos de librerias
                      cache = FALSE,    # cache de los chunks,usar analisis pesados
                      out.width = '85%',# largo de imagen en %
                      fig.pos= "H",     # posicion figuras H = HERE
                      echo = FALSE      # incluir chunk en output
                      )

@@

con <-c(
  "quintiles_ingresos_pc_factor",
  "libros_hogar_factor_ap",
  "univ",
  "pos_pol_factor_ap",
  "pos_pol_factor_est",
  "genero_factor_est",
  "dependencia_factor_ap"
  ) # Omitir coeficientes en la visualización

basereg$perc_tduro_bi_ap <- factor(basereg$perc_tduro_bi_ap,labels = c("Not important","Is important"))

# getah.or6 <- ordinal::clmm(formula(paste("perc_trabajo_duro_factor_est~",paste(c("perc_tduro_bi_ap","sj_indirect_ln","sj_direct_factor",con,ref),collapse="+"))), basereg, nAGQ = 10)
# 
# getah.or8 <- ordinal::clmm(formula(paste("perc_trabajo_duro_factor_est~",paste(c("perc_tduro_bi_ap","sj_indirect_ln","sj_direct_factor", con, "perc_tduro_bi_ap*sj_direct_factor","nse_grupo",ref),collapse="+"))), basereg)

getah.or6 <- lmer(formula(paste("as.numeric(perc_trabajo_duro_factor_est)~",paste(c("perc_tduro_bi_ap","sj_indirect_ln","sj_direct_factor","nse_grupo",con,ref),collapse="+"))), basereg)

getah.or8 <- lmer(formula(paste("as.numeric(perc_trabajo_duro_factor_est)~",paste(c("perc_tduro_bi_ap","sj_indirect_ln","sj_direct_factor", "perc_tduro_bi_ap*sj_direct_factor",con,ref),collapse="+"))), basereg)

rm_plot <- c(
  # "Not important|Somewhat important",
  # "Somewhat important|Important",
  # "Important|Very important",
  "quintiles_ingresos_pc_factorDon't know/No response",
  "libros_hogar_factor_apBetween 11 and 25 books",
  "libros_hogar_factor_apBetween 26 and 100 book",
  "libros_hogar_factor_apBetween 101 and 200 books",
  "libros_hogar_factor_apBetween 201 and 500 books",
  "libros_hogar_factor_apMore than 500 books",
  # "univ",
  "pos_pol_factor_apCenter Right",
  "pos_pol_factor_apCenter",
  "pos_pol_factor_apCenter Left",
  "pos_pol_factor_apLeft",
  "pos_pol_factor_apIndependent",
  "pos_pol_factor_apNone",
  "pos_pol_factor_apNa/Nr",
  "pos_pol_factor_estCenter Right",
  "pos_pol_factor_estCenter",
  "pos_pol_factor_estCenter Left",
  "pos_pol_factor_estLeft",
  "pos_pol_factor_estIndependent",
  "pos_pol_factor_estNone",
  "pos_pol_factor_estNa/Nr",
  "genero_factor_estMujer",
  "genero_factor_estOtro",
  "dependencia_factor_apPart. Subvencionado",
  "dependencia_factor_apPart. Privado")

@@

# plotgetah<- 
# plot_model(getah.or8,rm.terms = rm_plot,show.intercept = F,
#            show.p = T, show.values = T,
#            # order.terms = c(2,3,4,1,5,6,7,8,9,10,11,12),
#            axis.labels = rev(coef_names),
#            ) +
#   geom_hline(yintercept = 1)
# plotgetah

plotgetah<- 
plot_models(getah.or6,getah.or8,rm.terms = rm_plot,
            axis.labels = rev(coef_names),
            show.values = T,
            m.labels = c("Base", "Interactions"),
            legend.title = "Models") +
  geom_hline(yintercept = 0) +
  theme(legend.position = "bottom")

ggplot2::ggsave(plot = plotgetah,device = "png",units = "cm",
                width = 1,height = 1,scale = 18,
                path =  "docs/presentacion-effort/images/",
                filename = "plotgetahead.png")

# Elaborar modelo de interacción
# getah.or7 <- ordinal::clmm(formula(paste("perc_trabajo_duro_factor_est~",paste(c("perc_tduro_bi_ap","sj_indirect_noln","sj_direct_factor", con, "perc_tduro_bi_ap*sj_direct_factor",ref),collapse="+"))), basereg)
# 
# # Preparar datos para gráfico
# basereg$fitted_or7 <- getah.or7$fitted.values
# pred_or7<- basereg %>%
#   dplyr::mutate(sj_direct_factor=relevel(sj_direct_factor,ref = "Less than I deserve")) %>%
#   dplyr::group_by(perc_trabajo_duro_factor_est,perc_tduro_bi_ap,sj_direct_factor) %>%
#   dplyr::summarise(yhat=mean(fitted_or7))
# 
# # Elaborar gráfico interacción
# graph_interact <- ggplot2::ggplot(data=pred_or7, aes(x=perc_tduro_bi_ap, y=yhat,
#                           group=sj_direct_factor,
#                           color=sj_direct_factor,
#                          # fill=sj_direct_factor
#                          )) +
#   geom_point() +
#   geom_line()+
#   scale_x_discrete(labels = c("Not important", "Important")) +
#   facet_grid(~perc_trabajo_duro_factor_est) +
#   xlab("Importance of hard work (parents)") +
#   ylab("Probability") +
#   scale_color_discrete("Justice in grades - Direct") +
#   theme(legend.position="bottom")

graph_interact <- 
plot_model(getah.or8,type =  "int") +
  xlab("Importance of hard work (parents)") +
  ylab("Predicted values: Importance of hard work (students)") +
  ggtitle(NULL)+
  theme(legend.position = "bottom",legend.title = element_blank())

# 
# mtcars$gear <- factor(mtcars$gear)
# m_gear <- lm(mpg ~ gear * wt, data = mtcars)
# interplot(m = m_gear, var1 = "wt", var2 = "gear")
# 
# interplot::interplot(m = getah.or8, 
#                      var1 = "sj_direct_factor", 
#                      var2 = "perc_tduro_bi_ap")

# ggeffects::ggpredict(getah.or8,terms =c("perc_tduro_bi_ap","sj_direct_factor")) %>% plot()
# ggeffects::ggemmeans(getah.or8,terms =c("perc_tduro_bi_ap","sj_direct_factor")) %>% plot()

# graph_interact

ggplot2::ggsave(plot = graph_interact,device = "png",units = "cm",
                width = 1,height = 1,scale = 13,
                path =  "docs/presentacion-effort/images/",
                filename = "intergetahead.png")
```

@@

# effort.or6 <- ordinal::clmm(formula(paste("perc_esfuerzo_factor_est~",paste(c("perc_esfuerzo_bi_ap","sj_indirect_ln","sj_direct_factor", con,ref),collapse="+"))), basereg)
# effort.or8 <- ordinal::clmm(formula(paste("perc_esfuerzo_factor_est~",paste(c("perc_esfuerzo_bi_ap","sj_indirect_ln","sj_direct_factor", con, "perc_esfuerzo_bi_ap*sj_direct_factor","nse_grupo",ref),collapse="+"))), basereg)

effort.or6 <- lme4::lmer(formula(paste("as.numeric(perc_esfuerzo_factor_est)~",paste(c("perc_esfuerzo_bi_ap","sj_indirect_ln","sj_direct_factor", con,ref),collapse="+"))), basereg)
effort.or8 <- lme4::lmer(formula(paste("as.numeric(perc_esfuerzo_factor_est)~",paste(c("perc_esfuerzo_bi_ap","sj_indirect_ln","sj_direct_factor", con, "perc_esfuerzo_bi_ap*sj_direct_factor","nse_grupo",ref),collapse="+"))), basereg)

rm_plot <- c(
  # "Strongly Disagree|Disagree",
  # "Disagree|Agree",
  # "Agree|Strongly Agree",
  "quintiles_ingresos_pc_factorDon't know/No response",
  "libros_hogar_factor_apBetween 11 and 25 books",
  "libros_hogar_factor_apBetween 26 and 100 book",
  "libros_hogar_factor_apBetween 101 and 200 books",
  "libros_hogar_factor_apBetween 201 and 500 books",
  "libros_hogar_factor_apMore than 500 books",
  # "univ",
  "pos_pol_factor_apCenter Right",
  "pos_pol_factor_apCenter",
  "pos_pol_factor_apCenter Left",
  "pos_pol_factor_apLeft",
  "pos_pol_factor_apIndependent",
  "pos_pol_factor_apNone",
  "pos_pol_factor_apNa/Nr",
  "pos_pol_factor_estCenter Right",
  "pos_pol_factor_estCenter",
  "pos_pol_factor_estCenter Left",
  "pos_pol_factor_estLeft",
  "pos_pol_factor_estIndependent",
  "pos_pol_factor_estNone",
  "pos_pol_factor_estNa/Nr",
  "genero_factor_estMujer",
  "genero_factor_estOtro",
  "dependencia_factor_apPart. Subvencionado",
  "dependencia_factor_apPart. Privado")
